---
title: "Fairywren_Tutorial_10/28/25"
author: "Patrick Gritton"
format: html
editor: visual
---

## Qiime2 Fairy Wren Diet Analysis

## Introduction

This tutorial is a work in progress project notebook for my honor's thesis on fecal metabarcoding of fairy wren diet samples. I am working off of my previous rmarkdown projects from my microbiome analysis, the Qiime 2 moving picture tutorial, and Tasos Stephanou's "qiime_orchard_fecal_tutorial.md" tutorial.

This analysis will require: Rmarkdown, putty, filezilla, sublime, cisco secure client (WSL, Python)

-   Rmarkdown: For note taking and version control

-   putty: for connecting to the cluster

-   filezilla: for moving files on and off the cluster

-   sublime: for text editing

-   Cisco secure client: for enabling cornell VPN to have access to cluster outside of campus

-   (Optional): WSL and python for testing out the pipeline on a smaller subset of reads off of the cluster

## Lets get started

### Connecting to the VPN (ONLY NEEDED OFF CAMPUS)

1.  Open Cisco secure client

2.  type in cuvpn.cuvpn.cornell.edu

3.  click connect

### Connecting to the cluster

1.  Open Putty

2.  Fill in the following information

    1.  Host name (or IP address): cbsulogin.biohpc.cornell.edu

    2.  Port: 22

    3.  Connection type: SSH

3.  Click open

4.  Sign in

    1.  Username: RPG79

    2.  Password: Rpg79biohpc28503! (I need to make this shorter)

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
